---
title: "Live Session Unit 10 Assignment"
author: "M Nepal"
date: "July 21, 2017"
output:
  word_document: default
  pdf_document: default
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

## Introduction

This is R markdown document for keeping track of assignment submitted for <a href="https://datascience.smu.edu/">**MSDS-6306@SMU**</a> as an example of  **Exploratory Data Analysis (EDA)** based on dataset provided by Department of Statistics Columbia University in the City of New York.

There are 32 data sets named nyt0.csv, nyt1.csv,…, nyt31.csv, which can be downloaded from <a href="http://stat.columbia.edu/~rachel/datasets/">**here**.</a> 

Each csv represents one (simulated) days worth of ads shown and clicks recorded on the New York Times homepage in 2012. Each row in the csv represents a single user.

This information is taken from <a href="https://rpubs.com/tmcfl/simulated-click-analysis">RPubs</a>.

I have chosen <a href="http://stat.columbia.edu/~rachel/datasets/nyt1.csv">**nyt1.csv**</a> for EDA.

 
####Required packages
* ggplot2 
* plyr
* dplyr

Install and/or load these packages before trying the code below.

```{r}
library(ggplot2)
library(plyr)
library(dplyr)
```
```{r}
#Get the data from url
fileLocation <- "http://stat.columbia.edu/~rachel/datasets/nyt1.csv"
data1 <- read.csv(url(fileLocation))
names(data1) # This will help to know variable names.
```
```{r}
# str function provides the variable types.
str(data1)
```
###Exploratory Data Analysis
```{r}
#Let's find summary statistics of data set, just to make a good start for EDA
summary(data1)
```
####Create a new variable named ageGroup, that categorizes age into following groups:
####<18, 18-24, 25-34, 35-44, 45-54, 55-64, and 65+
```{r}
# categorizes age groups
head(data1)
data1$ageGroup <- cut(data1$Age, c(-Inf, 18, 24, 34, 44, 54, 64, Inf))
levels(data1$ageGroup) <- c("<18", "18-24", "25-34", "35-44", "45-54", "55-64", "65+")
summary(data1)
```

####Use sub set of data called "ImpSub" where Impressions > 0 
```{r}
ImpSub <- subset(data1, Impressions > 0) # new variable ImpSub
head(ImpSub)
```
####Create new variable called click-through-rate(CTR = click/impression)
```{r}
ImpSub$CTR <- ImpSub$Clicks/ImpSub$Impressions 
head(ImpSub)
```
####Plot distributions of number impressions and click-through-rate (CTR = click/impressions) for the age groups
```{r}
#Plot the distribution of Impressions>0, grouped by ageGroup
ggplot(subset(ImpSub, Impressions > 0), aes(x=Impressions, fill=ageGroup))+
  labs(title="Impressions by age group")+
  geom_histogram(binwidth=1)
```
```{r}
#Plot the distribution of CTR>0, grouped by ageGroup
ggplot(subset(ImpSub, CTR > 0), aes(x=CTR, fill=ageGroup))+
  labs(title="Click-through rate by age group")+
  geom_histogram(binwidth=0.25)
```

####Define a new variable to segment users based on click -through-rate (CTR) behavior.
####CTR< 0.2, 0.2 <= CTR < 0.4, 0.4 <= CTR < 0.6, 0.6 <= CTR <0.8, CTR >0.8

```{r}
ImpSub$CTR_Behavior <- cut(ImpSub$CTR, c(-Inf, 0.2, 0.4, 0.6, 0.8, Inf))
levels(ImpSub$CTR_Behavior) <- c("CTR < 0.2", "0.2 <= CTR < 0.4", "0.4 <= CTR < 0.6", "0.6 <= CTR < 0.8", "CTR  > 0.8")
```

#####7)	Get the total number of Male, Impressions, Clicks and Signed_In
#####(0=Female, 1=Male)

```{r}
str(ImpSub)
sapply(ImpSub[c(2,3,4,5)],sum) 
# The sum of gender also works here as Male=1 and Female=0, Gender represents total male since female = 0
```

####Get the mean of Age, Impressions, Clicks, CTR and percentage of males and signed_In 

```{r}

#Before
ImpSubPer <- sapply(ImpSub[c(1,3,4,7)],mean)
ImpSubPer
#Create percentage variables and combined with ImpSubPer
percentageOfMaleAndSigned_In <- c((sapply(ImpSub[c(2,5)],sum)/sapply(ImpSub[c(2,5)],length)*100))
percentageOfMaleAndSigned_In
ImpSubCombined <- c(ImpSubPer,percentageOfMaleAndSigned_In)
#after combining abd before cleaning col names
ImpSubCombined
##combined vector after cleaning the header for question 8
names(ImpSubCombined)<-c("Age_mean","Impressions_mean","Clicks_mean","CTR_mean","% of  Males","% of signed_in")
ImpSubCombined
```

####Get the means of Impressions, Clicks, CTR and percentage of males and signed_In  by AgeGroup.
```{r}
meansByAgeGroup <- aggregate(cbind(ImpSub$Impressions,ImpSub$Clicks,ImpSub$CTR)~ageGroup,FUN = mean,ImpSub,na.rm = TRUE)
colnames(meansByAgeGroup) <- c("ageGroup","Impressions_mean","Clicks_mean","CTR_mean")
meansByAgeGroup

#using dplyr/plyr package
sumOfMaleByAgeGroup <- ddply(ImpSub, "ageGroup", summarise, No_Of_Males=sum(Gender))
sumOfMaleByAgeGroup
sumOfSignedInAgeGroup <- ddply(ImpSub, "ageGroup", summarise, No_Of_Signed_In=sum(Signed_In))
#Incase you want to display
sumOfSignedInAgeGroup
combinedMaleandSign <- merge(sumOfMaleByAgeGroup,sumOfSignedInAgeGroup,by="ageGroup")
#In case if you want to display 
combinedMaleandSign
totalRows <- nrow(ImpSub)
totalRows

combinedMaleandSign$percentage_Of_Males <- ((combinedMaleandSign$No_Of_Males)/totalRows)*100
combinedMaleandSign
combinedMaleandSign$percentage_of_signed_In <- ((combinedMaleandSign$No_Of_Signed_In)/totalRows)*100
combinedMaleandSign
cleanedVector <- subset(combinedMaleandSign,select=c(1,4,5))# using dplyr package
cleanedVector
mergedvector <-merge(meansByAgeGroup, cleanedVector, by="ageGroup")
mergedvector
```
####Create a table of CTRGroup vs AgeGroup counts.
```{r}
ctr_age_Table <- table(ImpSub$CTR_Behavior,ImpSub$ageGroup)
ctr_age_Table
```
####Let's do One more plot
```{r}
hist(ImpSub$Age, main="Distribution of age", xlab="Age")
```


